---
title: "Part A, Topic 2 - Olympic Tweets"
author: "Lance Belen"
date: "`r Sys.Date()`"
output:
  pdf_document:
    includes:
      in_header: header.tex
    toc: true
    number_sections: true
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
# Data Wrangling, Visualisatsion, and Analysis
First, load the packages needed.
```{r load-packages, echo=TRUE, message=FALSE, warning=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org/"))

if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
if (!requireNamespace("lubridate", quietly = TRUE)) {
  install.packages("lubridate")
}
if (!requireNamespace("impute", quietly = TRUE)) {
  install.packages("impute")
}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
if (!requireNamespace("magrittr", quietly = TRUE)) {
  install.packages("magrittr")
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
if (!requireNamespace("readxl", quietly = TRUE)) {
  install.packages("readxl")
}
if (!requireNamespace("stringr", quietly = TRUE)) {
  install.packages("stringr")
}
if (!requireNamespace("forcats", quietly = TRUE)) {
  install.packages("forcats")
}
if (!requireNamespace("tidytext", quietly = TRUE)) {
  install.packages("tidytext")
}

library(tidyverse)
library(lubridate)
library(impute)
library(dplyr)
library(magrittr)
library(ggplot2)
library(readxl)
library(stringr)
library(forcats)
library(tidytext)
```

Then, read and open the dataset.
```{r read-datasets, echo=TRUE, message=FALSE, warning=FALSE}
tweets <- read.csv("Olympics_tweets.csv")
```

Next, explore the dataset.
```{r eval=FALSE}
cat("Number of rows: ", nrow(tweets), "\n")
cat("Number of columns: ", ncol(tweets), "\n")
knitr::kable(head(tweets))
cat("Structure of the dataset: \n")
str(tweets)
cat("Summary of the dataset: \n")
summary(tweets)
```
```{r tweets-data, echo=FALSE, message=TRUE, warning=FALSE}
cat("Number of rows: ", nrow(tweets), "\n")
cat("Number of columns: ", ncol(tweets), "\n")
knitr::kable(head(tweets))
cat("Structure of the dataset: \n")
str(tweets)
cat("Summary of the dataset: \n")
summary(tweets)
```

Now I will check the dataset for missing or incorrect values that needs handling.
```{r check-values, echo=TRUE, message=FALSE, warning=FALSE}
# retweet_count
tweets <- tweets %>%
  mutate(retweet_count = ifelse(retweet_count < 0, NA, retweet_count)) %>%
  mutate(retweet_count = replace_na(retweet_count, 0))

# favorite_count
tweets <- tweets %>%
  mutate(favorite_count = ifelse(favorite_count < 0, NA, favorite_count)) %>%
  mutate(favorite_count = replace_na(favorite_count, 0))

# user_followers
tweets <- tweets %>%
  mutate(user_followers = ifelse(user_followers < 0, NA, user_followers)) %>%
  mutate(user_followers = replace_na(user_followers, 0))

# user_friends
tweets <- tweets %>%
  mutate(user_friends = ifelse(user_friends < 0, NA, user_friends)) %>%
  mutate(user_friends = replace_na(user_friends, 0))

# user_created_at
tweets <- tweets %>%
  mutate(user_created_at = dmy_hm(user_created_at)) %>%
  filter(
    user_created_at >= as.POSIXct("2006-03-21"),
    user_created_at <= Sys.time()
  )

# date
tweets <- tweets %>%
  mutate(date = dmy_hm(date)) %>%
  filter(
    date >= as.POSIXct("2006-03-21"),
    date <= Sys.time()
  )
```

## Visualisation
### Number of Twitter Accounts Created per Year
```{r number-of-twitter-accounts, echo=TRUE, message=FALSE, warning=FALSE}
# Create a new column for the year
tweets$year <- year(tweets$user_created_at)
tweets$year <- as.integer(tweets$year)
tweets$year <- as.factor(tweets$year)

# Bar plot of the number of Twitter accounts per year
ggplot(tweets, aes(x = year)) +
  geom_bar(fill = "steelblue") +
  geom_text(
    stat = "count", aes(label = ..count..), 
    position = position_dodge(width = 0.9),
    vjust = -0.5) + 
  labs(x = "Year", y = "Number of Twitter Accounts Created") +
  theme_minimal()
```

### Average Followers per Year
```{r average-followers-per-year, echo=TRUE, message=FALSE, warning=FALSE}
creators_after_2010 <- tweets
creators_after_2010 <- creators_after_2010 %>% mutate(year = as.numeric(as.character(year)))
creators_after_2010 <- creators_after_2010 %>% filter(year > 2010)

creators_after_2010 %>%
  filter(year > 2010) %>%
  group_by(year) %>%
  summarise(avg_followers = mean(user_followers, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = avg_followers)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(
    aes(label = round(avg_followers, 2)), 
    vjust = -0.5) +
  scale_x_continuous(breaks = creators_after_2010$year) +
  labs(x = "Year", y = "Average Followers") +
  theme_minimal()
```

### Top 10 Most Frequest Location of Users
```{r top-10-most-frequent-location, echo=TRUE, message=FALSE, warning=FALSE}
top_locations <- tweets %>%
  select(user_screen_name, user_location) %>%
  distinct() %>%
  filter(!is.na(user_location), user_location != "") %>%
  count(user_location, name="user_count") %>%
  arrange(desc(user_count)) %>%
  slice(1:10)

top_locations <- top_locations %>%
  mutate(user_location = fct_reorder(user_location, user_count, .desc = TRUE)) %>%
  mutate(user_location = fct_rev(user_location))

ggplot(top_locations, aes(x = user_location, y = user_count)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = user_count), hjust = -0.1, vjust = 0.5) +
  coord_flip() +
  labs(x = "User Location", y = "Number of Users") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Number of Tweets Associated with the Top 10 locations
```{r number-of-tweets-associated-with-top-10-locations, echo=TRUE, message=FALSE, warning=FALSE}
tweets_in_top_10_locations <- tweets %>%
  filter(user_location %in% top_locations$user_location) %>%
  group_by(user_location) %>%
  summarise(tweet_count = n()) %>%
  arrange(desc(tweet_count))

tweets_in_top_10_locations <- tweets_in_top_10_locations %>%
  mutate(user_location = fct_reorder(user_location, tweet_count, .desc = TRUE)) %>%
  mutate(user_location = fct_rev(user_location))

ggplot(tweets_in_top_10_locations, aes(x = user_location, y = tweet_count)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = tweet_count), hjust = -0.1, vjust = 0.5) +
  coord_flip() +
  labs(x = "User Location", y = "Number of Tweets") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

From the last two visualisations, the odd values are "she/her", and some locations are the geographically the same from a broader perspective, but some are more specific to cities.
For example, "New York- NY", "Los Angeles- CA", and "United States". The first two locations are cities in the United States, but there is also
"United States" as a location.

### Bar Chart of the Length of Tweets
```{r bar-chart-of-length-of-tweets, echo=TRUE, message=FALSE, warning=FALSE}
tweets <- tweets %>%
  mutate(tweet_length = nchar(text)) %>%
  mutate(length_group = case_when(
    tweet_length >= 1 & tweet_length <= 40 ~ "1-40",
    tweet_length >= 41 & tweet_length <= 80 ~ "41-80",
    tweet_length >= 81 & tweet_length <= 120 ~ "81-120",
    tweet_length >= 121 & tweet_length <= 160 ~ "121-160",
    tweet_length >= 161 & tweet_length <= 200 ~ "161-200",
    tweet_length >= 201 & tweet_length <= 240 ~ "201-240",
    tweet_length >= 241 ~ "241+",
    TRUE ~ "Unknown"
  ))

length_group_counts <- tweets %>%
  count(length_group)

custom_order <- c("1-40", "41-80", "81-120", "121-160", "161-200", "201-240", "241+")
length_group_counts <- length_group_counts %>%
  mutate(length_group = factor(length_group, levels = custom_order))

ggplot(length_group_counts, aes(x = length_group, y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = n), vjust = -0.5) +
  labs(x = "Length of Tweets", y = "Number of Tweets") +
  theme_minimal()
```

### Number of Tweets with Mentions and Its Count
```{r number-of-tweets-with-mentions, echo=TRUE, message=FALSE, warning=FALSE}

tweets_with_mentions <- tweets %>%
  filter(str_detect(text, "@[A-Za-z0-9_]+")) %>%
  mutate(mentions_count = str_count(text, "@[A-Za-z0-9_]+"))

tweets_with_mentions <- tweets_with_mentions %>%
  group_by(mentions_count) %>%
  summarise(tweet_count = n()) %>%
  arrange(desc(tweet_count))

tweets_with_mentions <- tweets_with_mentions %>%
  mutate(mentions_count = as.factor(mentions_count)) %>%
  mutate(mentions_count = fct_reorder(mentions_count, tweet_count, .desc = TRUE)) %>%
  mutate(mentions_count = fct_rev(mentions_count))

ggplot(tweets_with_mentions, aes(x = mentions_count, y = tweet_count)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  geom_text(aes(label = tweet_count), hjust = -0.1, vjust = 0.5) +
  labs(x = "Number of Mentions", y = "Number of Tweets") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))

number_of_tweets_with_mentions <- tweets %>%
  filter(str_detect(text, "@[A-Za-z0-9_]+")) %>%
  nrow()

number_of_tweets_with_3_or_more_mentions <- tweets %>%
  filter(str_detect(text, "@[A-Za-z0-9_]+")) %>%
  mutate(mentions_count = str_count(text, "@[A-Za-z0-9_]+")) %>%
  filter(mentions_count >= 3) %>%
  nrow()

cat("Number of tweets with mentions: ", number_of_tweets_with_mentions, "\n")
cat("Number of tweets with 3 or more mentions: ", number_of_tweets_with_3_or_more_mentions, "\n")
```

### Top 20 Most Frequent Term in Tweets
```{r top-20-most-frequent-terms, echo=TRUE, message=FALSE, warning=FALSE}
library(tidytext)

tweets_frequent_terms <- tweets %>%
  mutate(text = tolower(text)) %>%
  unnest_tokens(word, text) %>%
  count(word, sort = TRUE)

top_20_terms <- tweets_frequent_terms %>%
  slice(1:20) %>%
  mutate(word = fct_reorder(word, n)) %>%
  mutate(word = fct_rev(word))

max_count <- max(top_20_terms$n)

ggplot(top_20_terms, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = n), hjust = -0.1, vjust = 0.5) +
  coord_flip() +
  labs(x = "Word", y = "Number of Occurrences") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Top 20 Most Frequent Non-Stopword Term in Tweets
```{r top-20-most-frequent-terms-with-no-stopwords, echo=TRUE, message=FALSE, warning=FALSE}
tweets_frequent_terms_no_stopwords <- tweets %>%
  mutate(text = tolower(text)) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE)

top_20_terms_no_stopwords <- tweets_frequent_terms_no_stopwords %>%
  slice(1:20) %>%
  mutate(word = fct_reorder(word, n)) %>%
  mutate(word = fct_rev(word))

max_count_no_stopwords <- max(top_20_terms_no_stopwords$n)

ggplot(top_20_terms_no_stopwords, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = n), hjust = -0.1, vjust = 0.5) +
  coord_flip() +
  labs(x = "Word", y = "Number of Occurrences") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```