---
title: "Part A, Topic 2 - Olympic Tweets"
author: |
  Lance Belen  
  Master of Applied Data Science  
  Monash University
date: "2 June 2025"

output:
  pdf_document:
    toc: true
    number_sections: true
    keep_tex: true
    latex_engine: xelatex
  html_document: default
  
nocite: '@*'
csl: https://www.zotero.org/styles/apa
references:
  - id: gradientboosting
    title: "What is Gradient Boosting?"
    author:
      - family: "Clark"
        given: "Bryan"
      - family: "Lee"
        given: "Fangfang"
    issued:
      - year: 2025
    url: "https://www.ibm.com/think/topics/gradient-boosting"
    type: webpage
  - id: xgboost
    title: "XGBoost: Everything You Need to Know"
    author:
      - family: "Hachcham"
        given: "Aymane"
    issued:
      - year: 2025
    url: "https://neptune.ai/blog/xgboost-everything-you-need-to-know"
    type: webpage
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\newpage

# Data Wrangling, Visualisatsion, and Analysis
First, load the packages needed.
```{r load-packages, echo=TRUE, message=FALSE, warning=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org/"))

if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
if (!requireNamespace("lubridate", quietly = TRUE)) {
  install.packages("lubridate")
}
if (!requireNamespace("impute", quietly = TRUE)) {
  install.packages("impute")
}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
if (!requireNamespace("magrittr", quietly = TRUE)) {
  install.packages("magrittr")
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
if (!requireNamespace("readxl", quietly = TRUE)) {
  install.packages("readxl")
}
if (!requireNamespace("stringr", quietly = TRUE)) {
  install.packages("stringr")
}
if (!requireNamespace("forcats", quietly = TRUE)) {
  install.packages("forcats")
}
if (!requireNamespace("tidytext", quietly = TRUE)) {
  install.packages("tidytext")
}
if(!requireNamespace("xgboost", quietly = TRUE)) {
  install.packages("xgboost")
}
if(!requireNamespace("MLmetrics", quietly = TRUE)) {
  install.packages("MLmetrics")
}

library(tidyverse)
library(lubridate)
library(impute)
library(dplyr)
library(magrittr)
library(ggplot2)
library(readxl)
library(stringr)
library(forcats)
library(tidytext)
library(xgboost)
library(MLmetrics)
```

## Data Wrangling
Then, read and open the dataset.
```{r read-datasets, echo=TRUE, message=FALSE, warning=FALSE}
tweets <- read.csv("Olympics_tweets.csv")
```

Next, explore the dataset.
```{r eval=FALSE}
cat("Number of rows: ", nrow(tweets), "\n")
cat("Number of columns: ", ncol(tweets), "\n")
knitr::kable(head(tweets))
cat("Structure of the dataset: \n")
str(tweets)
cat("Summary of the dataset: \n")
summary(tweets)
```
```{r tweets-data, echo=FALSE, message=TRUE, warning=FALSE}
cat("Number of rows: ", nrow(tweets), "\n")
cat("Number of columns: ", ncol(tweets), "\n")
knitr::kable(head(tweets))
cat("Structure of the dataset: \n")
str(tweets)
cat("Summary of the dataset: \n")
summary(tweets)
```

Now I will check the dataset for missing or incorrect values that needs handling.
```{r check-values, echo=TRUE, message=FALSE, warning=FALSE}
# retweet_count
tweets <- tweets %>%
  mutate(retweet_count = ifelse(retweet_count < 0, NA, retweet_count)) %>%
  mutate(retweet_count = replace_na(retweet_count, 0))

# favorite_count
tweets <- tweets %>%
  mutate(favorite_count = ifelse(favorite_count < 0, NA, favorite_count)) %>%
  mutate(favorite_count = replace_na(favorite_count, 0))

# user_followers
tweets <- tweets %>%
  mutate(user_followers = ifelse(user_followers < 0, NA, user_followers)) %>%
  mutate(user_followers = replace_na(user_followers, 0))

# user_friends
tweets <- tweets %>%
  mutate(user_friends = ifelse(user_friends < 0, NA, user_friends)) %>%
  mutate(user_friends = replace_na(user_friends, 0))

# user_created_at
tweets <- tweets %>%
  mutate(user_created_at = dmy_hm(user_created_at)) %>%
  filter(
    user_created_at >= as.POSIXct("2006-03-21"),
    user_created_at <= Sys.time()
  )

# date
tweets <- tweets %>%
  mutate(date = dmy_hm(date)) %>%
  filter(
    date >= as.POSIXct("2006-03-21"),
    date <= Sys.time()
  )
```

## Visualisations and Analysis
### Number of Twitter Accounts Created per Year
```{r number-of-twitter-accounts, echo=TRUE, message=FALSE, warning=FALSE}
# Create a new column for the year
tweets$year <- year(tweets$user_created_at)
tweets$year <- as.integer(tweets$year)
tweets$year <- as.factor(tweets$year)

# Bar plot of the number of Twitter accounts per year
ggplot(tweets, aes(x = year)) +
  geom_bar(fill = "steelblue") +
  geom_text(
    stat = "count", aes(label = ..count..), 
    position = position_dodge(width = 0.9),
    vjust = -0.5) + 
  labs(x = "Year", y = "Number of Twitter Accounts Created") +
  theme_minimal()
```

### Average Followers per Year
```{r average-followers-per-year, echo=TRUE, message=FALSE, warning=FALSE}
creators_after_2010 <- tweets
creators_after_2010 <- creators_after_2010 %>% mutate(year = as.numeric(as.character(year)))
creators_after_2010 <- creators_after_2010 %>% filter(year > 2010)

creators_after_2010 %>%
  filter(year > 2010) %>%
  group_by(year) %>%
  summarise(avg_followers = mean(user_followers, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = avg_followers)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(
    aes(label = round(avg_followers, 2)), 
    vjust = -0.5) +
  scale_x_continuous(breaks = creators_after_2010$year) +
  labs(x = "Year", y = "Average Followers") +
  theme_minimal()
```

### Top 10 Most Frequest Location of Users
```{r top-10-most-frequent-location, echo=TRUE, message=FALSE, warning=FALSE}
top_locations <- tweets %>%
  select(user_screen_name, user_location) %>%
  distinct() %>%
  filter(!is.na(user_location), user_location != "") %>%
  count(user_location, name="user_count") %>%
  arrange(desc(user_count)) %>%
  dplyr::slice(1:10)

top_locations <- top_locations %>%
  mutate(user_location = fct_reorder(user_location, user_count, .desc = TRUE)) %>%
  mutate(user_location = fct_rev(user_location))

ggplot(top_locations, aes(x = user_location, y = user_count)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = user_count), hjust = -0.1, vjust = 0.5) +
  coord_flip() +
  labs(x = "User Location", y = "Number of Users") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Number of Tweets Associated with the Top 10 locations
```{r number-of-tweets-associated-with-top-10-locations, echo=TRUE, message=FALSE, warning=FALSE}
tweets_in_top_10_locations <- tweets %>%
  filter(user_location %in% top_locations$user_location) %>%
  group_by(user_location) %>%
  summarise(tweet_count = n()) %>%
  arrange(desc(tweet_count))

tweets_in_top_10_locations <- tweets_in_top_10_locations %>%
  mutate(user_location = fct_reorder(user_location, tweet_count, .desc = TRUE)) %>%
  mutate(user_location = fct_rev(user_location))

ggplot(tweets_in_top_10_locations, aes(x = user_location, y = tweet_count)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = tweet_count), hjust = -0.1, vjust = 0.5) +
  coord_flip() +
  labs(x = "User Location", y = "Number of Tweets") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

From the last two visualisations, the odd values are "she/her", and some locations are the geographically the same from a broader perspective, but some are more specific to cities.
For example, "New York- NY", "Los Angeles- CA", and "United States". The first two locations are cities in the United States, but there is also
"United States" as a location.

### Bar Chart of the Length of Tweets
```{r bar-chart-of-length-of-tweets, echo=TRUE, message=FALSE, warning=FALSE}
tweets <- tweets %>%
  mutate(tweet_length = nchar(text)) %>%
  mutate(length_group = case_when(
    tweet_length >= 1 & tweet_length <= 40 ~ "1-40",
    tweet_length >= 41 & tweet_length <= 80 ~ "41-80",
    tweet_length >= 81 & tweet_length <= 120 ~ "81-120",
    tweet_length >= 121 & tweet_length <= 160 ~ "121-160",
    tweet_length >= 161 & tweet_length <= 200 ~ "161-200",
    tweet_length >= 201 & tweet_length <= 240 ~ "201-240",
    tweet_length >= 241 ~ "241+",
    TRUE ~ "Unknown"
  ))

length_group_counts <- tweets %>%
  count(length_group)

custom_order <- c("1-40", "41-80", "81-120", "121-160", "161-200", "201-240", "241+")
length_group_counts <- length_group_counts %>%
  mutate(length_group = factor(length_group, levels = custom_order))

ggplot(length_group_counts, aes(x = length_group, y = n)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = n), vjust = -0.5) +
  labs(x = "Length of Tweets", y = "Number of Tweets") +
  theme_minimal()
```

### Number of Tweets with Mentions and Its Count
```{r number-of-tweets-with-mentions, echo=TRUE, message=FALSE, warning=FALSE}

tweets_with_mentions <- tweets %>%
  filter(str_detect(text, "@[A-Za-z0-9_]+")) %>%
  mutate(mentions_count = str_count(text, "@[A-Za-z0-9_]+"))

tweets_with_mentions <- tweets_with_mentions %>%
  group_by(mentions_count) %>%
  summarise(tweet_count = n()) %>%
  arrange(desc(tweet_count))

tweets_with_mentions <- tweets_with_mentions %>%
  mutate(mentions_count = as.factor(mentions_count)) %>%
  mutate(mentions_count = fct_reorder(mentions_count, tweet_count, .desc = TRUE)) %>%
  mutate(mentions_count = fct_rev(mentions_count))

ggplot(tweets_with_mentions, aes(x = mentions_count, y = tweet_count)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  geom_text(aes(label = tweet_count), hjust = -0.1, vjust = 0.5) +
  labs(x = "Number of Mentions", y = "Number of Tweets") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))

number_of_tweets_with_mentions <- tweets %>%
  filter(str_detect(text, "@[A-Za-z0-9_]+")) %>%
  nrow()

number_of_tweets_with_3_or_more_mentions <- tweets %>%
  filter(str_detect(text, "@[A-Za-z0-9_]+")) %>%
  mutate(mentions_count = str_count(text, "@[A-Za-z0-9_]+")) %>%
  filter(mentions_count >= 3) %>%
  nrow()

cat("Number of tweets with mentions: ", number_of_tweets_with_mentions, "\n")
cat("Number of tweets with 3 or more mentions: ", number_of_tweets_with_3_or_more_mentions, "\n")
```

### Top 20 Most Frequent Term in Tweets
```{r top-20-most-frequent-terms, echo=TRUE, message=FALSE, warning=FALSE}
library(tidytext)

tweets_frequent_terms <- tweets %>%
  mutate(text = tolower(text)) %>%
  unnest_tokens(word, text) %>%
  count(word, sort = TRUE)

top_20_terms <- tweets_frequent_terms %>%
  dplyr::slice(1:20) %>%
  mutate(word = fct_reorder(word, n)) %>%
  mutate(word = fct_rev(word))

max_count <- max(top_20_terms$n)

ggplot(top_20_terms, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = n), hjust = -0.1, vjust = 0.5) +
  coord_flip() +
  labs(x = "Word", y = "Number of Occurrences") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```

### Top 20 Most Frequent Non-Stopword Term in Tweets
```{r top-20-most-frequent-terms-with-no-stopwords, echo=TRUE, message=FALSE, warning=FALSE}
tweets_frequent_terms_no_stopwords <- tweets %>%
  mutate(text = tolower(text)) %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE)

top_20_terms_no_stopwords <- tweets_frequent_terms_no_stopwords %>%
  dplyr::slice(1:20) %>%
  mutate(word = fct_reorder(word, n)) %>%
  mutate(word = fct_rev(word))

max_count_no_stopwords <- max(top_20_terms_no_stopwords$n)

ggplot(top_20_terms_no_stopwords, aes(x = reorder(word, n), y = n)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = n), hjust = -0.1, vjust = 0.5) +
  coord_flip() +
  labs(x = "Word", y = "Number of Occurrences") +
  theme_minimal() +
  scale_y_continuous(expand = expansion(mult = c(0, 0.15)))
```
\vspace{2cm}

# Predictive Data Analysis
## Model Training
In this section, I will build machine learning models to predict whether a tweet is relevant or not.
I will use the glm function to build a logistic regression models using all of the features from the dataset.
The glm function is used to fit generalized linear models, which can be used for binary classification tasks like this one.
Next, I will try to use interaction terms to see if that can improve the same model's performance.
FInally, I will use the xgboost package to build a gradient boosting model to also evaluate how it performs compared to the logistic regression models.

First, I will make sure that the dependent variable is numeric and has only two values, 0 and 1.
Luckily, the dataset has already been split into training and testing sets, so I will use the training set to build the models and the testing set to evaluate them.
```{r prepare-data-for-modeling, echo=TRUE, message=FALSE, warning=FALSE}
# Load the training and testing datasets
train <- read.csv("predictive_twitter_train.csv")

# Check missing values in relevanceJudge column
cat("Missing values in relevanceJudge column: ", 
    sum(is.na(train$relevanceJudge)), "\n")

# Check if relevanceJudge is numeric
cat("Is relevanceJudge numeric: ", 
    is.numeric(train$relevanceJudge), "\n")

# Number of columns in the training set
cat("Number of columns in the training set: ", 
    ncol(train), "\n")

# Number of columns that are numeric in the training set
cat("Number of numeric columns in the training set: ", 
    sum(sapply(train, is.numeric)), "\n")

# Exclude tweetID from the training set
train <- train %>%
  select(-tweetID)

# Rename mistakenly named column nFavorties to nFavorites
train <- train %>%
  rename(nFavorites = nFavorties)
```

From the output in the code chunk above, we can see that the `relevanceJudge` column has no missing values, is numeric, and there are 26 columns in the training set, all of which are numeric.
Next, I will build a logistic regression model using the `glm` function. I will use the `relevanceJudge` column as the dependent variable and all other columns as independent variables.

```{r build-logistic-regression-model, echo=TRUE, message=FALSE, warning=FALSE}
# Build the logistic regression model
model1 <- glm(relevanceJudge ~ ., data = train, family = binomial)

# Summary of the model
summary(model1)

# Check the coefficients of the model
cat("Coefficients of the model with all of the features: \n")
print(coef(model1))
```

Next, I will build another model with interaction terms between some of the author's account details.
```{r build-logistic-regression-model-with-interaction-terms, echo=TRUE, message=FALSE, warning=FALSE}
# Build the logistic regression model with interaction terms
model2 <- glm(relevanceJudge ~ isVerified *
              nFollowers *
              nFriends *
              nFavorites *
              nListed +
              .,
              data = train, family = binomial)

# Summary of the model
summary(model2)

# Check the coefficients of the model
cat("Coefficients of the model with interaction terms:\n")
print(coef(model2))
```

Finally, I will build an XGboost model using the `xgboost` package.
```{r build-logistic-xgboost, echo=TRUE, message=FALSE, warning=FALSE}
# Prepare the data for XGBoost
train_clean <- na.omit(train)  # Remove rows with NA values
train_matrix <- model.matrix(relevanceJudge ~ ., data = train_clean)[, -1]
train_label <- train_clean$relevanceJudge

# Build the XGBoost model
model3 <- xgboost(data = train_matrix, label = train_label, 
                  nrounds = 100, objective = "binary:logistic",
                  eval_metric = "logloss", verbose = 0)

# Check the importance of the features
cat("Feature importance of the XGBoost model: \n")
print(xgb.importance(model = model3))
```

## Model Evalutaion
Now I will evaluate the performance of the models by measuring their F1 score.
```{r evaluate-models, echo=TRUE, message=FALSE, warning=FALSE}
# Make predictions on the train set using the models just to evaluate their performance.
# I will use the train set later for the final model's predictions.
predictions_model1 <- predict(model1, newdata = train, type = "response")
predictions_model2 <- predict(model2, newdata = train, type = "response")
predictions_model3 <- predict(model3, newdata = train_matrix)

# Convert predictions to binary values (0 or 1) using a threshold of 0.5
predictions_model1_binary <- ifelse(predictions_model1 > 0.5, 1, 0)
predictions_model2_binary <- ifelse(predictions_model2 > 0.5, 1, 0)
predictions_model3_binary <- ifelse(predictions_model3 > 0.5, 1, 0)

# Get the F1 score for each model
f1_score_model1 <- F1_Score(y_true = train$relevanceJudge, y_pred = predictions_model1_binary)
f1_score_model2 <- F1_Score(y_true = train$relevanceJudge, y_pred = predictions_model2_binary)
f1_score_model3 <- F1_Score(y_true = train_label, y_pred = predictions_model3_binary)
```
```{r eval=FALSE}
# Print the F1 scores
cat("F1 Score for Model 1 (Logistic Regression): ", f1_score_model1, "\n")
cat("F1 Score for Model 2 (Logistic Regression with interaction terms): ", f1_score_model2, "\n")
cat("F1 Score for Model 3 (XGBoost Model): ", f1_score_model3, "\n")
```
```{r f1-scores, echo=FALSE, message=FALSE, warning=FALSE}
# Print the F1 scores
cat("F1 Score for Model 1 (Logistic Regression): ", f1_score_model1, "\n")
cat("F1 Score for Model 2 (Logistic Regression with interaction terms): ", f1_score_model2, "\n")
cat("F1 Score for Model 3 (XGBoost Model): ", f1_score_model3, "\n")
```

From the output of the code chunk above, we can see that the first and the second model that used logistic regression have the same F1 score rounded up to 2 decimal places - 0.96.
The third model that used XGBoost performed better than the preceding two models with an F1 score of 0.99, rounded up to 2 decimal places.

## Final Model Improvements
I will now try to improve the XGBoost model further, which has an exact F1 score of 0.9854909.
I will filter out any outliers in the dataset, deal with missing values, and select the most important features based on the feature importance from the XGBoost model.
```{r final-model-improvements, echo=TRUE, message=FALSE, warning=FALSE}
# Number of rows before filtering outliers and handling missing values
cat("Number of rows before filtering outliers: ", nrow(train), "\n")

# Columns to omit rows with NA values
columns_for_na_omit <- c("length", "nFollowers", "nFriends", "isVerified")

# Columns to calculate mean for missing values
columns_for_mean <- c("text_score", "text_score_expansion",
  "twitterAge", "X.tweetsPosted")

# Columns to get mode for missing values
columns_for_mode <- c("nFavorites", "nListed")

# Handle missing values by getting either omitting the row, calculating the mean, or getting the mode of the column
train_filtered <- train %>%
  filter(if_all(all_of(columns_for_na_omit), ~ !is.na(.)))

train_filtered <- train_filtered %>%
  mutate(across(all_of(columns_for_mean),
                ~ ifelse(is.na(.), mean(., na.rm = TRUE), .)))

train_filtered <- train_filtered %>%
  mutate(across(all_of(columns_for_mode),
                ~ ifelse(is.na(.), as.numeric(names(sort(table(.), decreasing = TRUE)[1])), .)))

# Number of rows after handling missing values
cat("Number of rows after handling missing values: ", nrow(train_filtered), "\n")

# Select columns to filter outliers
columns_to_filter <- c("text_score", "text_score_expansion",
  "length", "nFollowers", "nFriends",
  "nFavorites", "nListed", "twitterAge", "X.tweetsPosted")

# Filter out outliers based on the interquartile range (IQR)
train_filtered <- train_filtered %>%
  filter(across(all_of(columns_to_filter), ~ . >= quantile(., 0.05) & . <= quantile(., 0.95)))

# Number of rows after filtering outliers
cat("Number of rows after filtering outliers: ", nrow(train_filtered), "\n")
```

Now I will build the final XGBoost model using the filtered dataset and the most important features from the previous XGBoost model.
```{r build-final-xgboost-model, echo=TRUE, message=FALSE, warning=FALSE}
# Prepare the data for XGBoost
importance_matrix <- xgb.importance(model = model3)

# Select the 10 most important features
top_features <- importance_matrix$Feature[1:10]
train_matrix_final <- model.matrix(relevanceJudge ~ ., data = train_filtered[, c(top_features, "relevanceJudge")])[, -1]
train_label_final <- train_filtered$relevanceJudge

# Build the final XGBoost model
model_final <- xgboost(data = train_matrix_final, label = train_label_final, 
                       nrounds = 100, objective = "binary:logistic",
                       eval_metric = "logloss", verbose = 0)

# Check the importance of the features in the final model
cat("Feature importance of the final XGBoost model: \n")
print(xgb.importance(model = model_final))

# Check the F1 score of the final model
predictions_final <- predict(model_final, newdata = train_matrix_final)
predictions_final_binary <- ifelse(predictions_final > 0.5, 1, 0)
f1_score_final <- F1_Score(y_true = train_label_final, y_pred = predictions_final_binary)
cat("F1 Score for the final XGBoost model: ", f1_score_final, "\n")
```

By handling missing values and filtering out outliers, the final XGBoost model has an exact F1 score of 0.9880673, which is a 0.0025764 improvement from the previous XGBoost model's F1 score of 0.9854909.

### Justifications of the Final Model
As per [@gradientboosting], "Gradient boosting is an ensemble learning algorithm that produces accurate predictions by combining multiple decision trees into a single model."
Essentially, the algorithm builds a series of decision trees that corrects the errors of the preceding trees, which makes it a powerful model for complex datasets.
I figured that this type of algorithm is particularly useful for this dataset because a decision tree applies for such classification tasks.

Hence, I've decided to use XGBoost model as one of the three models that I built because it is a powerful implementation of gradient boosting that can handle datasets of all sizes and complex relationships between features.
It also has built-in regularization to prevent overfitting, which is important for this dataset with many features. Given that this is a provided dataset and I do not have any prior knowledge of the data, XGBoost
was a great choice as it does well in handling datasets where the relationship between predictors and the outcome is complex. (@xgboost)

As mentioned earlier, the initial XGBoost model already had an F1 score of 0.99, rounded up to 2 decimal places.
There wasn't much room for improvement, but I was able to improve the F1 score by 0.0025764 by filtering out outliers and handling missing values.
I chose follow that method because it is a common yet a reliably effective way to improve the performance of a model, even if it already has an accuracy of close to 100%.
And since I haven't performed those preprocessing steps yet, I thought it was the best choice to improve the model's performance.

## Predictions on the Testing Set
Finally, I will use the final XGBoost model to make predictions on the testing set and save the results in its corresponding CSV file.
```{r predict-on-testing-set, echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)

original_test <- read.csv("predictive_twitter_test.csv")
test <- original_test %>%
  rename(nFavorites = nFavorties)

for (col in top_features) {
  if (is.numeric(test[[col]])) {
    test[[col]][is.na(test[[col]])] <- mean(test[[col]], na.rm = TRUE)
  }
}

# Make model matrix for XGBoost and predict
test_matrix <- model.matrix(~ . - 1, data = test[, top_features])
predictions <- predict(model_final, newdata = test_matrix)
predictions_binary <- ifelse(predictions > 0.5, 1, 0)

original_test$relevanceJudge <- predictions_binary
write.csv(original_test, "predictive_twitter_test.csv", row.names = FALSE)

cat("Number of relevant tweets in the test set: ", 
    sum(original_test$relevanceJudge == 1), "\n")

cat("Number of irrelevant tweets in the test set: ", 
    sum(original_test$relevanceJudge == 0), "\n")
```
\newpage

# References