---
title: "Part A, Topic 1 - COVID-19"
author: "Lance Belen"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document:
    includes:
      in_header: header.tex
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
# Data Wrangling and Integration

First, load the packages needed.
```{r load-packages, echo=TRUE, message=FALSE, warning=FALSE}
options(repos = c(CRAN = "https://cloud.r-project.org/"))

if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}
if (!requireNamespace("lubridate", quietly = TRUE)) {
  install.packages("lubridate")
}
if (!requireNamespace("impute", quietly = TRUE)) {
  install.packages("impute")
}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
if (!requireNamespace("magrittr", quietly = TRUE)) {
  install.packages("magrittr")
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
if (!requireNamespace("readxl", quietly = TRUE)) {
  install.packages("readxl")
}
if (!requireNamespace("stringr", quietly = TRUE)) {
  install.packages("stringr")
}
if (!requireNamespace("forcats", quietly = TRUE)) {
  install.packages("forcats")
}
if (!requireNamespace("tidytext", quietly = TRUE)) {
  install.packages("tidytext")
}

library(tidyverse)
library(lubridate)
library(impute)
library(dplyr)
library(magrittr)
library(ggplot2)
library(readxl)
library(stringr)
library(forcats)
library(tidytext)
```

Then, read and open the three datasets, Covid-data.csv, CountryLockdowndates.csv and WorldwideVaccineData.csv
```{r read-datasets, echo=TRUE, message=FALSE, warning=FALSE}
covid <- read.csv("Covid-data.csv")
lockdown <- read.csv("CountryLockdowndates.csv")
vaccine <- read.csv("WorldwideVaccineData.csv")
```

## Exploration
**Covid-data.csv**
```{r, eval=FALSE}
cat("Number of rows:", nrow(covid), "\n")
cat("Number of columns:", ncol(covid), "\n")
cat("Number of missing values:", sum(is.na(covid)), "\n")
knitr::kable(head(covid), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(covid)
cat("Summary of the dataset:\n")
summary(covid)
```
```{r covid-data, echo=FALSE, message=TRUE, warning=FALSE}
cat("Number of rows:", nrow(covid), "\n")
cat("Number of columns:", ncol(covid), "\n")
cat("Number of missing values:", sum(is.na(covid)), "\n")
knitr::kable(head(covid), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(covid)
cat("Summary of the dataset:\n")
summary(covid)
```
\vspace{1cm}

**CountryLockdowndates.csv**
```{r eval=FALSE}
cat("Number of rows:", nrow(lockdown), "\n")
cat("Number of columns:", ncol(lockdown), "\n")
cat("Number of missing values:", sum(is.na(lockdown)), "\n")
knitr::kable(head(lockdown), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(lockdown)
cat("Summary of the dataset:\n")
summary(lockdown)
```
```{r lockdown-data, echo=FALSE, message=TRUE, warning=FALSE}
cat("Number of rows:", nrow(lockdown), "\n")
cat("Number of columns:", ncol(lockdown), "\n")
cat("Number of missing values:", sum(is.na(lockdown)), "\n")
knitr::kable(head(lockdown), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(lockdown)
cat("Summary of the dataset:\n")
summary(lockdown)
```
\vspace{1cm}

**WorldwideVaccineData.csv**
```{r eval=FALSE}
cat("Number of rows:", nrow(vaccine), "\n")
cat("Number of columns:", ncol(vaccine), "\n")
cat("Number of missing values:", sum(is.na(vaccine)), "\n")
knitr::kable(head(vaccine), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(vaccine)
summary(vaccine)
```
```{r vaccine-data, echo=FALSE, message=TRUE, warning=FALSE}
cat("Number of rows:", nrow(vaccine), "\n")
cat("Number of columns:", ncol(vaccine), "\n")
cat("Number of missing values:", sum(is.na(vaccine)), "\n")
knitr::kable(head(vaccine), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(vaccine)
summary(vaccine)
```
\vspace{2cm}

## Wrangling
I have chosen to use KNN imputation to deal with missing values or wrong values such as negative values for new cases, deaths, vaccinated, etc.
From the exploration above, only covid has missing values, so I will impute that dataset using K Nearest Neighbors (KNN).
By using KNN, I can fill in the missing values based on the values of the nearest neighbors in the dataset, allowing us keep the rest of the data for those entries.
```{r eval=FALSE}
library(impute)
library(dplyr)
library(magrittr)

cat("Number of missing values before imputation:", sum(is.na(covid)), "\n")

covid_imputed <- covid %>%
  select(where(is.numeric)) %>%
  as.matrix() %>%
  impute.knn(k = 5) %>%
  .$data %>%
  as.data.frame()

covid[, colnames(covid_imputed)] <- covid_imputed

cat("Number of missing values after imputation:", sum(is.na(covid)), "\n")
```
```{r knn-imputation, echo=FALSE, message=FALSE, warning=FALSE}
library(impute)
library(dplyr)
library(magrittr)

cat("Number of missing values before imputation:", sum(is.na(covid)), "\n")

covid_imputed <- covid %>%
  select(where(is.numeric)) %>%
  as.matrix() %>%
  impute.knn(k = 5) %>%
  .$data %>%
  as.data.frame()

covid[, colnames(covid_imputed)] <- covid_imputed

cat("Number of missing values after imputation:", sum(is.na(covid)), "\n")
```
\vspace{1cm}

I also want to remove any negative values from the dataset as they are not valid in some of the available features.
First, check if there are indeed any entries that have such values.\newline
\vspace{1cm}

**Covid-data.csv**
```{r check-negative-values-covid, echo=FALSE, message=FALSE, warning=FALSE}
covid %>%
  select(where(is.numeric)) %>%
  summarise(across(everything(), ~ any(. < 0))) %>%
  print()
```

In this case, new_cases and new_deaths are the only columns that have negative values, but those values are not valid.
I will remove any rows that have negative values in those columns.
```{r eval=FALSE}
cat("Number of rows before removing negative values:", nrow(covid), "\n")
covid <- covid %>%
  filter(new_cases >= 0 & new_deaths >= 0)
cat("Number of rows after removing negative values:", nrow(covid), "\n")
```
```{r remove-negative-values-covid, echo=FALSE, message=FALSE, warning=FALSE}
cat("Number of rows before removing negative values:", nrow(covid), "\n")
covid <- covid %>%
  filter(new_cases >= 0 & new_deaths >= 0)
cat("Number of rows after removing negative values:", nrow(covid), "\n")
```
Since CountryLockdowndates.csv doesn't have numerical values, I will now proceed to WorldwideVaccineData.csv.\newline
\vspace{1cm}

**WorldwideVaccineData.csv**
```{r check-negative-values-vaccine, echo=FALSE, message=FALSE, warning=FALSE}
vaccine %>%
  select(where(is.numeric)) %>%
  summarise(across(everything(), ~ any(. < 0))) %>%
  print()
```
With this, I can confirm that there are no negative values in this dataset, so no further action is needed.
\vspace{1cm}

## Integration
Now that I have cleaned the datasets, we can integrate them into one. Firstly, the date attributes are in different formats for Covid-data.csv and CountryLockdowndates.csv, so I need to convert them to the same format.
```{r eval=FALSE}
covid$date <- as.Date(covid$date, format = "%Y-%m-%d")
lockdown$Date <- as.Date(lockdown$Date, format = "%d/%m/%Y")

cat("Covid-data.csv date format: ", class(covid$date), "\nHead: ", head(covid$date))
cat("CountryLockdowndates.csv date format: ", class(lockdown$Date), "\nHead: ", head(lockdown$Date))
```
```{r date-formatting, echo=FALSE, message=FALSE, warning=FALSE}
covid$date <- as.Date(covid$date, format = "%Y-%m-%d")
lockdown$Date <- as.Date(lockdown$Date, format = "%d/%m/%Y")

cat("Covid-data.csv date format: ", class(covid$date), "\nHead: ", head(covid$date))
cat("CountryLockdowndates.csv date format: ", class(lockdown$Date), "\nHead: ", head(lockdown$Date))
```
\vspace{1cm}

Next, I will rename the columns in the datasets as per the requirements of the merged dataset.\newline
The expected columns are:\newline
- location\newline
- date\newline
- total_cases\newline
- new_cases\newline
- total_deaths\newline
- new_deaths\newline
- gdp_per_capita\newline
- population\newline
- lockdown_date\newline
- total doses administered\newline
- % of population fully vaccinated\newline
``` {r rename-columns, echo=TRUE, message=FALSE, warning=FALSE}
lockdown <- lockdown %>%
  rename(location = Country.Region,
         lockdown_date = Date)
vaccine <- vaccine %>%  
  rename(location = Country,
         `total doses administered` = `Total.doses.administered`,
         `% of population fully vaccinated` = `X..of.population.fully.vaccinated`)
```
\vspace{1cm}

Now I can join the three datasets together and review any further preprocessing that needs to be done.\newline
I will use left_join to keep all the rows from the covid dataset where the country is also present in the other two datasets.
Based on the visualisation and analysis that will be conducted later, there is no need to keep entries in the other two datasets where the country is not present in the covid dataset.
```{r join-datasets, echo=TRUE, message=FALSE, warning=FALSE}

data <- covid %>%
  left_join(lockdown %>% select(location, lockdown_date), by = "location") %>%
  left_join(vaccine %>% select(location, `total doses administered`, `% of population fully vaccinated`), by = "location")
```
\vspace{1cm}

```{r eval=FALSE}
cat("Number of rows:", nrow(data), "\n")
cat("Number of columns:", ncol(data), "\n")
cat("Number of missing values:", sum(is.na(covid %>% select(where(is.numeric)))), "\n")
knitr::kable(head(data), caption = "First 6 rows of the merged dataset")
cat("Structure of the merged dataset:\n")
str(data)
cat("Summary of the merged dataset:\n")
summary(data)
```
```{r review-merged-dataset, echo=FALSE, message=FALSE, warning=FALSE}
cat("Number of rows:", nrow(data), "\n")
cat("Number of columns:", ncol(data), "\n")
cat("Number of missing values:", sum(is.na(covid %>% select(where(is.numeric)))), "\n")
knitr::kable(head(data), caption = "First 6 rows of the merged dataset")
cat("Structure of the merged dataset:\n")
str(data)
cat("Summary of the merged dataset:\n")
summary(data)
```
```{r eval=FALSE}
unique(data$location)
```
``` {r message=TRUE}
unique(data$location)
```
I can see that some of the values in the location column are not consistent, so I will need to clean them up accordingly.
```{r tidy-location, echo=TRUE, message=FALSE, warning=FALSE}
data$location <- data$location %>%
  str_replace_all("iran", "Iran") %>%
  str_replace_all("Itly", "Italy") %>%
  str_replace_all("UnitedKingdom", "United Kingdom") %>%
  str_replace_all("United Stats", "United States") %>%
  str_trim()
```

After cleaning up:
```{r eval=FALSE}
unique(data$location)
```
``` {r message=TRUE}
unique(data$location)
```

\vspace {2 cm}

# Data Visualisation and Analysis
``` {r line-plot, echo=TRUE, message=FALSE, warning=FALSE}
library(scales)

data %>%
  filter(!is.na(new_cases), !is.na(date), !is.na(location)) %>%
  ggplot(aes(x = date, y = new_cases, color = location)) +
  geom_line(size = 1, alpha = 0.5) +
  labs(title = "Trend of New COVID-19 Cases by Country", x = "Date", y = "New Cases") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  theme_minimal() +
  theme(
    legend.position = "bottom", legend.title = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 10)),
    axis.title.y = element_text(size = 12, face = "bold", margin = margin(r = 10)),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5, margin = margin(b = 20)),
  ) +
  scale_y_continuous(labels = comma)
```

The first occurrence of a notable observation of new cases is in China,
which reflects the fact that it is the origin of the novel coronavirus outbreak,
also known as COVID-19 (Centers for Disease Control and Prevention, 2024).
A significant increase can also be seen in the United States, compared to other countries in the dataset,
which started after about a month since China’s spike.
This event forced a declaration in the United States for a nationwide emergency and further overseas
travel bans on the 13th of March, 2020. (Centers for Disease Control and Prevention, 2024).
By August 2020, the reported total number of cases in U.S.A. was 5.04 million, which is evident of the
prolonged large number of new cases daily as seen in the plot (Bergquist, S., Otten, T., & Sarich, N., 2020).

Apart from the two significant findings previously mentioned, it can also be seen that the trend of the number
of new cases are similar among the European countries. By February 2020, the virus has reached Europe, with Italy
being the first to react with lockdowns in different provinces as cases began to rise. By April 2020, numerous
countries in Europe have reached 10,000 total cases, which is also evident in the plot (Chadwick, L., 2020)

The observed shift in the increase of new cases from China to other countries supports 
the fact that the outbreak originated in China. As international travel occurs daily,
the spread of the virus to other places was inevitable, which explains the gradual and fluctuating rise in cases.
Because of this, numerous countries implemented lockdowns and travel bans to control the spread of the virus,
leading to a slow decline in new cases, which became more apparent by the middle to late part of 2020 in most countries.
\vspace{1cm}


Now here is an observation of the relationship between the death rate and new case numbers with the GDP of each country.
```{r mean-gdp, echo=TRUE, message=FALSE, warning=FALSE}
mean_gdp <- data %>%
  summarise(mean_gdp = mean(gdp_per_capita, na.rm = TRUE)) %>%
  pull(mean_gdp)
cat("Mean GDP: ", mean_gdp, "\n")
```
```{r eval=FALSE}
data <- data %>%
  mutate(GDP_Status = ifelse(gdp_per_capita >= mean_gdp, 1, 0))

cat("Number of rows with GDP Status 1: ", nrow(data[data$GDP_Status == 1, ]), "\n")
cat("Number of rows with GDP Status 0: ", nrow(data[data$GDP_Status == 0, ]), "\n")
```
```{r col-gdp-status, echo=FALSE, message=FALSE, warning=FALSE}
data <- data %>%
  mutate(GDP_Status = ifelse(gdp_per_capita >= mean_gdp, 1, 0))

cat("Number of rows with GDP Status 1 (GDP greater than or equal to the mean): ", nrow(data[data$GDP_Status == 1, ]), "\n")
cat("Number of rows with GDP Status 0 (GDP less than the mean): ", nrow(data[data$GDP_Status == 0, ]), "\n")
```
```{r eval=FALSE}
data <- data %>%
  mutate(new_case_rate = new_cases / population,
         new_death_rate = new_deaths / population)

head(data %>% filter(new_case_rate > 0 & new_death_rate > 0), 10)
```
```{r new-case-death-rate, echo=FALSE, message=FALSE, warning=FALSE}
data <- data %>%
  mutate(new_case_rate = new_cases / population,
         new_death_rate = new_deaths / population)

head(data %>% filter(new_case_rate > 0 & new_death_rate > 0), 10)
```

```{r scatter-plot-new-case-death-rate, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data, aes(x = new_case_rate, y = new_death_rate, color = factor(GDP_Status))) +
  geom_point(alpha = 0.7) +
  scale_color_manual(
    values = c("0" = "firebrick", "1" = "steelblue"),
    labels = c("0" = "Less than average", "1" = "Equal to or greater than average"),
    name = "GDP Category"
  ) +
  labs(
    title = "Daily Case and Death Rate by GDP Status",
    x = "Daily Case Rate",
    y = "Daily Death Rate"
  ) +
  theme_minimal()
```

The Gross Domestic Product, or GDP, is a representation of the monetary value of goods and services produced by a country in a
specified period. (Callen, T., 2025) If this value is to be divided by the total population of a country, we would get the GDP per capita,
which is the number that we have in our dataset that represents a rough measure of a country’s living standards, or simply, the
financial capability of a person in the country.

From the scatter plot above, it can be seen that countries with lower than average GDP per capita tend to have lower daily
new/death case rates. This can be due to multiple factors such as the country lacking the capability to have robust healthcare
and data systems, leading to underreporting of the number of new cases and deaths per day. It can also be linked to factors such
as the life expectancy and tourist spending being significantly lower than those that have average or greater GDP per capita.
To provide an example:

- Luxembourg is identified as having the highest GDP per capita. Burundi is identified as having the lowest estimated GDP per capita. (The World Factbook, 2023).\newline
- Life expectancy in Luxembourg is 82.8 years. Burundi’s is 64 years. Almost 20 years difference (World Health Organisation, 2023).\newline
- Luxembourg tourist spending for 2017 was 5.47 billion US Dollars (macrotrends, 2025). Burundi tourist spending for 2017 was 0.1 billion US dollars. (Knoema Data Appliance, 2025).

Overseas travel played a major role in the global spread of COVID-19 as rapid transmission across borders occurred. In 2020, the death rate from the virus
for those that are aged 85 and over was 1,645 per 100,000 population -- seven times greater than the rate for those aged 65 to 74. (Tejada-Vera, B. & Kramarow, A., 2022).

# Predictive Data Analaysis