---
title: "Part A, Topic 1 - COVID-19"
author: "Lance Belen"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document:
    includes:
      in_header: header.tex
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

\newpage
# Data Wrangling and Integration

First, load the packages needed.
```{r load-packages, echo=TRUE, message=FALSE, warning=FALSE}
library(readxl)
library(ggplot2)
library(tidyverse)
```

Then, read and open the three datasets, Covid-data.csv, CountryLockdowndates.csv and WorldwideVaccineData.csv
```{r read-datasets, echo=TRUE, message=FALSE, warning=FALSE}
covid <- read.csv("Covid-data.csv")
lockdown <- read.csv("CountryLockdowndates.csv")
vaccine <- read.csv("WorldwideVaccineData.csv")
```

## Exploration
**Covid-data.csv**
```{r, eval=FALSE}
cat("Number of rows:", nrow(covid), "\n")
cat("Number of columns:", ncol(covid), "\n")
cat("Number of missing values:", sum(is.na(covid)), "\n")
knitr::kable(head(covid), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(covid)
cat("Summary of the dataset:\n")
summary(covid)
```
```{r covid-data, echo=FALSE, message=TRUE, warning=FALSE}
cat("Number of rows:", nrow(covid), "\n")
cat("Number of columns:", ncol(covid), "\n")
cat("Number of missing values:", sum(is.na(covid)), "\n")
knitr::kable(head(covid), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(covid)
cat("Summary of the dataset:\n")
summary(covid)
```
\vspace{1cm}

**CountryLockdowndates.csv**
```{r eval=FALSE}
cat("Number of rows:", nrow(lockdown), "\n")
cat("Number of columns:", ncol(lockdown), "\n")
cat("Number of missing values:", sum(is.na(lockdown)), "\n")
knitr::kable(head(lockdown), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(lockdown)
cat("Summary of the dataset:\n")
summary(lockdown)
```
```{r lockdown-data, echo=FALSE, message=TRUE, warning=FALSE}
cat("Number of rows:", nrow(lockdown), "\n")
cat("Number of columns:", ncol(lockdown), "\n")
cat("Number of missing values:", sum(is.na(lockdown)), "\n")
knitr::kable(head(lockdown), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(lockdown)
cat("Summary of the dataset:\n")
summary(lockdown)
```
\vspace{1cm}

**WorldwideVaccineData.csv**
```{r eval=FALSE}
cat("Number of rows:", nrow(vaccine), "\n")
cat("Number of columns:", ncol(vaccine), "\n")
cat("Number of missing values:", sum(is.na(vaccine)), "\n")
knitr::kable(head(vaccine), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(vaccine)
summary(vaccine)
```
```{r vaccine-data, echo=FALSE, message=TRUE, warning=FALSE}
cat("Number of rows:", nrow(vaccine), "\n")
cat("Number of columns:", ncol(vaccine), "\n")
cat("Number of missing values:", sum(is.na(vaccine)), "\n")
knitr::kable(head(vaccine), caption = "First 6 rows of the dataset")
cat("Structure of the dataset:\n")
str(vaccine)
summary(vaccine)
```
\vspace{2cm}

## Wrangling
I have chosen to use KNN imputation to deal with missing values or wrong values such as negative values for new cases, deaths, vaccinated, etc.
From the exploration above, only covid has missing values, so I will impute that dataset using K Nearest Neighbors (KNN).
By using KNN, I can fill in the missing values based on the values of the nearest neighbors in the dataset, allowing us keep the rest of the data for those entries.
```{r eval=FALSE}
library(impute)
library(dplyr)
library(magrittr)

cat("Number of missing values before imputation:", sum(is.na(covid)), "\n")

covid_imputed <- covid %>%
  select(where(is.numeric)) %>%
  as.matrix() %>%
  impute.knn(k = 5) %>%
  .$data %>%
  as.data.frame()

covid[, colnames(covid_imputed)] <- covid_imputed

cat("Number of missing values after imputation:", sum(is.na(covid)), "\n")
```
```{r knn-imputation, echo=FALSE, message=FALSE, warning=FALSE}
library(impute)
library(dplyr)
library(magrittr)

cat("Number of missing values before imputation:", sum(is.na(covid)), "\n")

covid_imputed <- covid %>%
  select(where(is.numeric)) %>%
  as.matrix() %>%
  impute.knn(k = 5) %>%
  .$data %>%
  as.data.frame()

covid[, colnames(covid_imputed)] <- covid_imputed

cat("Number of missing values after imputation:", sum(is.na(covid)), "\n")
```
\vspace{1cm}

I also want to remove any negative values from the dataset as they are not valid in some of the available features.
First, check if there are indeed any entries that have such values.\newline
\vspace{1cm}

**Covid-data.csv**
```{r check-negative-values-covid, echo=FALSE, message=FALSE, warning=FALSE}
covid %>%
  select(where(is.numeric)) %>%
  summarise(across(everything(), ~ any(. < 0))) %>%
  print()
```

In this case, new_cases and new_deaths are the only columns that have negative values, but those values are not valid.
I will remove any rows that have negative values in those columns.
```{r eval=FALSE}
cat("Number of rows before removing negative values:", nrow(covid), "\n")
covid <- covid %>%
  filter(new_cases >= 0 & new_deaths >= 0)
cat("Number of rows after removing negative values:", nrow(covid), "\n")
```
```{r remove-negative-values-covid, echo=FALSE, message=FALSE, warning=FALSE}
cat("Number of rows before removing negative values:", nrow(covid), "\n")
covid <- covid %>%
  filter(new_cases >= 0 & new_deaths >= 0)
cat("Number of rows after removing negative values:", nrow(covid), "\n")
```
Since CountryLockdowndates.csv doesn't have numerical values, I will now proceed to WorldwideVaccineData.csv.\newline
\vspace{1cm}

**WorldwideVaccineData.csv**
```{r check-negative-values-vaccine, echo=FALSE, message=FALSE, warning=FALSE}
vaccine %>%
  select(where(is.numeric)) %>%
  summarise(across(everything(), ~ any(. < 0))) %>%
  print()
```
With this, I can confirm that there are no negative values in this dataset, so no further action is needed.
\vspace{1cm}

## Integration
Now that I have cleaned the datasets, we can integrate them into one. Firstly, the date attributes are in different formats for Covid-data.csv and CountryLockdowndates.csv, so I need to convert them to the same format.
```{r eval=FALSE}
covid$date <- as.Date(covid$date, format = "%Y-%m-%d")
lockdown$Date <- as.Date(lockdown$Date, format = "%d/%m/%Y")

cat("Covid-data.csv date format: ", class(covid$date), "\nHead: ", head(covid$date))
cat("CountryLockdowndates.csv date format: ", class(lockdown$Date), "\nHead: ", head(lockdown$Date))
```
```{r date-formatting, echo=FALSE, message=FALSE, warning=FALSE}
covid$date <- as.Date(covid$date, format = "%Y-%m-%d")
lockdown$Date <- as.Date(lockdown$Date, format = "%d/%m/%Y")

cat("Covid-data.csv date format: ", class(covid$date), "\nHead: ", head(covid$date))
cat("CountryLockdowndates.csv date format: ", class(lockdown$Date), "\nHead: ", head(lockdown$Date))
```
\vspace{1cm}

Next, I will rename the columns in the datasets as per the requirements of the merged dataset.\newline
The expected columns are:\newline
- location\newline
- date\newline
- total_cases\newline
- new_cases\newline
- total_deaths\newline
- new_deaths\newline
- gdp_per_capita\newline
- population\newline
- lockdown_date\newline
- total doses administered\newline
- % of population fully vaccinated\newline
``` {r rename-columns, echo=TRUE, message=FALSE, warning=FALSE}
lockdown <- lockdown %>%
  rename(location = Country.Region,
         lockdown_date = Date)
vaccine <- vaccine %>%  
  rename(location = Country,
         `total doses administered` = `Total.doses.administered`,
         `% of population fully vaccinated` = `X..of.population.fully.vaccinated`)
```
\vspace{1cm}

Now I can join the three datasets together and review any further preprocessing that needs to be done.\newline
I will use left_join to keep all the rows from the covid dataset where the country is also present in the other two datasets.
Based on the visualisation and analysis that will be conducted later, there is no need to keep entries in the other two datasets where the country is not present in the covid dataset.
```{r join-datasets, echo=TRUE, message=FALSE, warning=FALSE}

data <- covid %>%
  left_join(lockdown %>% select(location, lockdown_date), by = "location") %>%
  left_join(vaccine %>% select(location, `total doses administered`, `% of population fully vaccinated`), by = "location")
```
\vspace{1cm}

```{r eval=FALSE}
cat("Number of rows:", nrow(data), "\n")
cat("Number of columns:", ncol(data), "\n")
cat("Number of missing values:", sum(is.na(covid %>% select(where(is.numeric)))), "\n")
knitr::kable(head(data), caption = "First 6 rows of the merged dataset")
cat("Structure of the merged dataset:\n")
str(data)
cat("Summary of the merged dataset:\n")
summary(data)
```
```{r review-merged-dataset, echo=FALSE, message=FALSE, warning=FALSE}
cat("Number of rows:", nrow(data), "\n")
cat("Number of columns:", ncol(data), "\n")
cat("Number of missing values:", sum(is.na(covid %>% select(where(is.numeric)))), "\n")
knitr::kable(head(data), caption = "First 6 rows of the merged dataset")
cat("Structure of the merged dataset:\n")
str(data)
cat("Summary of the merged dataset:\n")
summary(data)
```
```{r eval=FALSE}
unique(data$location)
```
``` {r message=TRUE}
unique(data$location)
```
I can see that some of the values in the location column are not consistent, so I will need to clean them up accordingly.
```{r tidy-location, echo=TRUE, message=FALSE, warning=FALSE}
data$location <- data$location %>%
  str_replace_all("iran", "Iran") %>%
  str_replace_all("Itly", "Italy") %>%
  str_replace_all("UnitedKingdom", "United Kingdom") %>%
  str_replace_all("United Stats", "United States") %>%
  str_trim()
```

After cleaning up:
```{r eval=FALSE}
unique(data$location)
```
``` {r message=TRUE}
unique(data$location)
```

\vspace {2 cm}

# Data Visualisation and Analysis
``` {r line-plot, echo=TRUE, message=FALSE, warning=FALSE}
library(scales)

data %>%
  filter(!is.na(new_cases), !is.na(date), !is.na(location)) %>%
  ggplot(aes(x = date, y = new_cases, color = location)) +
  geom_line(size = 1, alpha = 0.5) +
  labs(title = "Trend of New COVID-19 Cases by Country", x = "Date", y = "New Cases") +
  scale_x_date(date_labels = "%b %Y", date_breaks = "1 month") +
  theme_minimal() +
  theme(
    legend.position = "bottom", legend.title = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 10)),
    axis.title.y = element_text(size = 12, face = "bold", margin = margin(r = 10)),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5, margin = margin(b = 20)),
  ) +
  scale_y_continuous(labels = comma)
```

The line plot shows significant increase in new cases in the United States compared to other countries in the dataset.\newline
The other country that is notable is Chine, as it showed the first spike in new cases, followed by Italy and Spain.\newline
TODO: events surrounding significatant points on the plots (cause of the spike, or sharp reduction in cases, etc.).\newline
The trend is similar for Spain, Italy and France, followed closely by Iran.\newline
TODO: my understanding of the trend.
\vspace{1cm}

Now I will observe the relationship between the death rate and new case numbers with the GDP of each country.
```{r mean-gdp, echo=TRUE, message=FALSE, warning=FALSE}
mean_gdp <- data %>%
  summarise(mean_gdp = mean(gdp_per_capita, na.rm = TRUE)) %>%
  pull(mean_gdp)
cat("Mean GDP: ", mean_gdp, "\n")
```
```{r eval=FALSE}
data <- data %>%
  mutate(GDP_Status = ifelse(gdp_per_capita >= mean_gdp, 1, 0))

cat("Number of rows with GDP Status 1: ", nrow(data[data$GDP_Status == 1, ]), "\n")
cat("Number of rows with GDP Status 0: ", nrow(data[data$GDP_Status == 0, ]), "\n")
```
```{r col-gdp-status, echo=FALSE, message=FALSE, warning=FALSE}
data <- data %>%
  mutate(GDP_Status = ifelse(gdp_per_capita >= mean_gdp, 1, 0))

cat("Number of rows with GDP Status 1 (GDP greater than or equal to the mean): ", nrow(data[data$GDP_Status == 1, ]), "\n")
cat("Number of rows with GDP Status 0 (GDP less than the mean): ", nrow(data[data$GDP_Status == 0, ]), "\n")
```
```{r eval=FALSE}
data <- data %>%
  mutate(new_case_rate = new_cases / population,
         new_death_rate = new_deaths / population)

head(data %>% filter(new_case_rate > 0 & new_death_rate > 0), 10)
```
```{r new-case-death-rate, echo=FALSE, message=FALSE, warning=FALSE}
data <- data %>%
  mutate(new_case_rate = new_cases / population,
         new_death_rate = new_deaths / population)

head(data %>% filter(new_case_rate > 0 & new_death_rate > 0), 10)
```

```{r scatter-plot-new-case-death-rate, echo=TRUE, message=FALSE, warning=FALSE}
ggplot(data, aes(x = new_case_rate, y = new_death_rate, color = factor(GDP_Status))) +
  geom_point(alpha = 0.7) +
  scale_color_manual(
    values = c("0" = "firebrick", "1" = "steelblue"),
    labels = c("0" = "Less than average", "1" = "Equal to or greater than average"),
    name = "GDP Category"
  ) +
  labs(
    title = "Daily Case and Death Rate by GDP Status",
    x = "Daily Case Rate",
    y = "Daily Death Rate"
  ) +
  theme_minimal()
```